---
title: "Discussion Week 2 - Data 605"
author: "Stephanie Roark"
date: "2/6/2019"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
library(devtools)
install_github("markheckmann/mat2tex")
```

###Linear Algebra Problem DM.C28

Doing the computations by hand, find the determinant of the matrix A.

Definition DM Determinant of a Matrix Suppose A is a square matrix. 

$$
\begin{bmatrix} 
a & b & c \\
d & e & f \\ 
g & h & i \\ 
\end{bmatrix}
$$

Then its determinant is:

A = a(ei - fh) - b(di - fg) + c(dh - eg)


First let's find the determinant using the det() function.

```{r determinant}
A <- matrix(c(1,0,1,1,2,-1,-1,1,2,5,3,0,1,-1,0,1), nrow = 4, byrow = TRUE)
print(A)
det(A)
round(det(A))
```
 
And now by hand:

The pattern for 4×4 matrices:

- plus a times the determinant of the matrix that is not in a's row or column,
- minus b times the determinant of the matrix that is not in b's row or column,
- plus c times the determinant of the matrix that is not in c's row or column,
- minus d times the determinant of the matrix that is not in d's row or column.

```{r byhand, echo=FALSE}


submatrix1_A <- matrix(c(-1,-1,1,5,3,0,-1,0,1), nrow = 3, byrow = TRUE)
submatrix2_A <- 0
submatrix3_A <- matrix(c(2,-1,1,2,3,0,1,0,1), nrow = 3, byrow = TRUE)
submatrix4_A <- matrix(c(2,-1,-1,2,5,3,1,-1,0), nrow = 3, byrow = TRUE)
```

Now to calculate the determinant:

$$
\begin{bmatrix} 
1 & 0 & 1 & 1\\
2 & -1 & -1 & 1 \\ 
2 & 5 & 3 & 0 \\ 
1 & -1 & 0 & 1 \\
\end{bmatrix}
=
\begin{bmatrix} 
-1 & -1 & 1 \\ 
5 & 3 & 0 \\ 
-1 & 0 & 1 \\
\end{bmatrix}
- 0
+
\begin{bmatrix} 
2 & -1 & 1 \\ 
2 & 5 & 0 \\ 
1 & -1 & 1 \\
\end{bmatrix}
-
\begin{bmatrix} 
2 & -1 & -1 \\ 
2 & 5 & 3 \\ 
1 & -1 & 0 \\
\end{bmatrix}
$$


```{r det}
detA = (1*(-1*(3*1 - 0*0) - -1*(5*1 - 0*-1) + 1*(5*0 - 3*-1)) 
        - 0 
        +1*(2*(5*1 - 0*-1) - -1*(2*1 - 0*1) + 1*(2*-1 - 5*1))
        -1*( 2*(5*0 - 3*-1)- -1*(2*0 - 3*1) + -1*(2*-1 - 5*1)))
detA
```
