---
title: "Data 605 Assignment 2"
author: "Stephanie Roark"
date: "2/10/2019"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    highlight: pygments
    theme: cosmo
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```

##1. Problem Set 1

####(1) Show that A^T^A != AA^T^ in general. (Proof and demonstration.)

If A is an m × n matrix and A^T^ is its transpose, then the result of matrix multiplication with these two matrices gives two square matrices: AA^T^ is m × m and A^T^A is n × n. Since an mxm matrix cannot be equavilent to an nxn matrix if m!=n, A^T^A != AA^T^.

```{r transpose A}
A <- matrix(c(1,2,3,4,5,6,7,8,9),nrow = 3)
A

A_T <- t(A)
A_T

identical(A, A_T)

A %*% A_T == A_T %*% A
```

####(2) For a special type of square matrix A, we get A^T^A = AA^T^. Under what conditions could this be true? 

#####The Identity matrix I is an example of such a matrix.
 
```{r transpose AI}
AI <- matrix(c(1,0,0,0,1,0,0,0,1),nrow = 3)
AI

AI_T <- t(AI)
AI_T

identical(AI, AI_T)

AI %*% AI_T == AI_T %*% AI 
```


#####This will also be true whenever A = A^T^.

```{r AT2}
A2 <- matrix(c(1,3,4,3,2,3,4,3,1), nrow = 3)
A2

A2_T <- t(A2)
A2_T

identical(A2, A2_T)

A2 %*% A2_T == A2_T %*% A2 
```

##2. Problem set 2

Matrix factorization is a very important problem. There are supercomputers built just
to do matrix factorizations. Every second you are on an airplane, matrices are being
factorized. Radars that track flights use a technique called Kalman filtering. At the heart
of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear
systems of equations when they track your flight using radars.

####Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer. 

Factorize A into a product of two matrices: A = LU where U is the matrix that we get at the end of the elimination procedure, it is the Upper Triangular matrix. L is a Lower Triangular matrix and you’ll see that the entries of L are the multipliers that we applied to subtract one row from the other.

#####LU Decomposition

```{r LU}
#define a matrix a
a <- matrix(sample (25), 5, 5)
a
#define Gaussian Elimination function to compute matrix with row reductions
Gauss_Elim <- function(a) {
    n <- nrow(a)
    #create an augmented matrix with a and the identity
    a <- cbind(a,diag(n))
    
    #multiply a row by a constant    
    for (i in 1 : n) {
        c <- diag(n)
        c[i, i] <- (1 / a[i, i])
        a <- c %*% a
       
        #if at the last row, stop
        if (i == n) {
            break ()
        }
        #add a multiple of a row to a different row
        for (j in (i + 1) : n) {
            r <- diag(n)
            r[j, i] <- (- a[j, i])
            a <- r %*% a    
        }
    }
    return(a)
}
#call Gauss Elimination function on matrix a
Gauss_Elim(a)

#use Gauss_Elim() to compute Upper Triangular Matrix and Lower Triangular Matrix

#define LU Decomposition function
LU_decomp <- function(a) {
    n <- nrow (a)
    #compute the U matrix (Upper Triangular)
    g <- Gauss_Elim(a)
    #compute the L matrix (Lower Triangular) with multipliers located below the diagonal 
    h <- Gauss_Elim(g[, n + 1 : n])
    return(list(L = h[, n + 1 : n], U = g[, 1 : n]))
}

print(LU <- LU_decomp(a))
#show that the LU decomp is equal to matrix a
LU $ L %*% LU $ U
```


