---
title: "Data607Homework2"
author: "Stephanie Roark"
date: "9/9/2018"
output:
  html_document:
    theme: cosmo
    highlight: pygments
---

#Assignment – SQL and R

Choose six recent popular movies. Ask at least five people that you know (friends, family, classmates,
imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5. Take the results
(observations) and store them in a SQL database. Load the information into an R dataframe.

###Reading from the Database

Connecting to MySql and reading in the reviews and movies tables from the movies database.

```{r get password prompt, eval=TRUE, echo=FALSE}
#Rmd will not knit with this password prompt, but code works stand alone. Add your password in place of dbPassword.
dbPassword <- rstudioapi::askForPassword()
movie_db = dbConnect(MySQL(), user='root', password=dbPassword, dbname='Movies', host='localhost')

```

```{r read in the tables - database,   eval=TRUE }
library(DBI)
library(RMySQL)

movie_db = dbConnect(MySQL(), user='root', password=dbPassword, dbname='Movies', host='localhost')

reviews <- dbGetQuery(movie_db, "SELECT * FROM reviewers ORDER BY movie_ID, UsersName")

movies <- dbGetQuery(movie_db, "SELECT movies.Title, AVG(movies.Rotten_Rating), AVG(reviewers.Reviewer_Rating) FROM movies INNER JOIN  reviewers ON movies.movie_ID = reviewers.movie_ID GROUP BY movies.Title")
```

##Reviews Table

Each person gave a rating and a review for each of the 10 recent popular movies.

```{r reviews table, eval=TRUE}
str(reviews)
colnames(reviews) <- c('MovieID', 'ReviewerName', 'Rating', 'Review')
reviews
```

##Movies Tables

Each movie is listed with the Rotten Tomatoes average rating and the sample reviewers average ratings.

```{R movies table, eval=TRUE}
str(movies)
colnames(movies) <- c('Title', 'RottenTomatesRating', 'AvgReviewerRating')
movies

```

## Comparing Rotten Tomatoes Ratings to Sample Ratings

We plot the ratings side by side to compare the Rotten Tomatoes ratings with our sample reviewers ratings.

```{r change col names and plot, eval=TRUE}

colours = c("red","green")
par(mar=c(5,9,4,4))
barplot(t(as.matrix(movies[2:3])), 
        main="Movie Ratings", xlab = "Average Rating",
        names.arg = movies$Title, horiz = TRUE, las = 1,
        cex.lab = 1, cex.main = 1.4, 
        cex.names = 0.5,
        beside=TRUE, col=colours)
par(xpd=TRUE)
legend(-2,-4, c("Rotten Tomatoes","Personal Reviewer"), 
       cex=0.8, bty="n", fill=colours)
```
