---
title: "Data 607 Homework 3"
author: "Stephanie Roark"
date: "9/16/2018"
output:output:
  html_document:
    theme: cosmo
    highlight: pygments
---

## Extracting the names from the Simpson's Phone Directory


```{r data}
raw.data <-("555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert")
```

## Question 3
### a) Use the tools of this chapter to rearrange the vector so that all the elements conform to the standard first_name last_name.

```{r extract the names, eval=TRUE}
library(stringr)
name <- str_extract_all(raw.data, "[[:alpha:]., ]{2,}")
name
# Check for ",". If no comma, is first_name last_name, if yes comma, reverse to make first_name last-name
reverse <- function(full_name) {
    fl_name <- ifelse(str_detect(full_name,","),
                      paste( str_split(full_name, ", ")[[1]][[2]], str_split(full_name, ", ")[[1]][[1]] ),
                      full_name)
    str_split( str_replace(fl_name, "[[:alpha:]]{1,}\\. ", ""), " ")
}
Simpson_names <- lapply(name[[1]], reverse)
Simpson_names
```

### b) Construct a logical vector indicating whether a character has a title.
```{r has title?, eval=TRUE}
has_title <- function(full_name) !is.na(str_extract(full_name, "[[:alpha:]]{2,}\\."))
name
lapply(name, has_title)
```

### c) Construct a logical vector indicating whether a character has a second name
```{r has second name?, eval=TRUE}
has_secname <- function(full_name) !is.na(str_extract(full_name, "\\b[[:alpha:]]{1,1}\\."))
has_secname <- function(full_name) !is.na(str_extract(full_name, "\\b[[:alpha:]]{1,1}\\."))
name
lapply(name, has_secname)
```

## Question 4
### Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

### a) [0-9]+\\$

One or more numbers followed by a dollar sign

```{r}
str_detect("12345$", "[0-9]+\\$")
```

### b) \\b[a-z]{1,4}\\b

A lower case character string that is between 1 to 4 letters long

```{r}
str_detect("asgf","\\b[a-z]{1,4}\\b") # TRUE
```

### c) .*?\\.txt$

0 or more (ungreedy) digits of any character followed by a . and txt with nothing after the txt
In other words a name of a text file ending .txt

```{r}
str_detect("jad34.txt", ".*?\\.txt$") # TRUE
```

### d) \\d{2}/\\d{2}/\\d{4}

2 digits / 2 digits/ 4 digits
which would match a date such as 09/16/2018

```{r}
str_detect("01/01/2018","\\d{2}/\\d{2}/\\d{4}")
```

### e) <(.+?)>.+?</\\1>

A character string of any length between <> and </> as in an html tag. 
<html>I really love regex!!!!!!!!!!!</html>

```{r}
str_detect("<html>I really love regex!!!!!!!!!!!</html>","<(.+?)>.+?</\\1>")
```





