---
title: "Homework12"
author: "Stephanie Roark"
date: "11/25/2018"
params:
  dbPassword: "to be completed interactively"
output:
  html_document:
    theme: cosmo
---

##MySQL to NoSQL

For this assignment, you should take information from a relational database and migrate it to a NoSQL database of your own choosing. 

###Flights

We begin by loading the flights database into R from MySQL by creating a connection and downloading the tables.

```{r MySQL, eval=TRUE}
library(DBI)
library(RMySQL)
library(mongolite)
library(knitr)

flights_db = dbConnect(MySQL(), user='root', password= params$dbPassword, dbname='flights', host='localhost')
dbListTables(flights_db)

flights<-dbGetQuery(flights_db,"SELECT * FROM flights;")
airlines<-dbGetQuery(flights_db,"SELECT * FROM airlines;")
airports<-dbGetQuery(flights_db,"SELECT * FROM airports;")
planes<-dbGetQuery(flights_db,"SELECT * FROM planes;")
weather<-dbGetQuery(flights_db,"SELECT * FROM weather;")

dbDisconnect(flights_db)
```

Next we load the tables into data frames in R.

```{r dfTables, eval=TRUE}
myflights <- data.frame(flights)
myairlines <- data.frame(airlines)
myairports <- data.frame(airports)
myplanes <- data.frame(planes)
myweather <- data.frame(weather)

kable(head(myweather))
```

Now we can load these data frames into a MongoDB database using the mongolite package in R.

```{r mongo, eval=TRUE}
mongodb <- mongo(collection = "myflights", db = "flights_db", url = "mongodb://localhost",
      verbose = FALSE, options = ssl_options())

mongodb$insert(myflights)

mongodb <- mongo(collection = "myairlines", db = "flights_db", url = "mongodb://localhost",
                 verbose = FALSE, options = ssl_options())

mongodb$insert(myairlines)

mongodb <- mongo(collection = "myairports", db = "flights_db", url = "mongodb://localhost",
                 verbose = FALSE, options = ssl_options())

mongodb$insert(myairports)

mongodb <- mongo(collection = "myplanes", db = "flights_db", url = "mongodb://localhost",
                 verbose = FALSE, options = ssl_options())

mongodb$insert(myplanes)

mongodb <- mongo(collection = "myweather", db = "flights_db", url = "mongodb://localhost",
                 verbose = FALSE, options = ssl_options())

mongodb$insert(myweather)
```

And finally we query the mongodb data base to show that the tables where loaded the MondoDB.

```{r mongodb_query, eval=TRUE}
mongoflights <- mongodb$find(limit = 5)
mongoflights

```

### Compare relational databases with NoSQL databases

A relational database like MySQL uses tables which can be linked and queried using a structured query language. SQL databases are vertically scalable which allows for easliy increasing storage capacity. Some benefits of relational databases are maturity, compatibility, replicability, and cost effectivenes. 

NoSQL databases are not table based, but rather store the data in column-oriented, document-oriented, graph-based or organized as a KeyValue store and therefore are more flexible in their format. The language used to query is a dynamic schema for unstructured data. A NoSQL database like MongoDB for example stores data in json form. NoSQL databases are horizontally scalable which allows for more traffic. Some benefits are scalability, manageablity, speed, and flexibility.